---
title: "rWDPA example to query GBIF"
author: "Luis J. Villanueva-Rivera"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignettes works as an example of using the `rWDPA` to search for a protected area and then get the occurrence reports from GBIF for that area.

## Get a list of countries in the WDPA database

```{r, eval = FALSE}
countries <- getWDPACountries(filename = 'countries.csv', api_key='xxxxxxxxxxx')
```

From that list, we see that the ISO code for Puerto Rico is 'PRI'. Use this ISO code in the other functions.

## Search for a protected areas in a country

```{r, eval = FALSE}
areas <- getWDPACountryAreas(country_iso = 'PRI', filename = 'areas.csv', api_key='xxxxxxxxxxx')
```

From this list, the WDPA ID of the protected area 'Isla de Mona' is 12438. 

## Get details of a protected area

Use the WDPA ID to get the details of a protected area, including the geometry.

```{r, eval = FALSE}
islamona <- getWDPA(wdpaid = 12438, with_geometry = TRUE, filename = 'islamona.geojson', api_key='xxxxxxxxxxx')
```

## Get occurrence reports from GBIF for a protected area

With the geometry, we can now query GBIF for species reports. This example uses the packages `wellknown` and `rgbif`.

```{r, eval = FALSE}
library(rgbif)
library(wellknown)

#Save the geometry as WKT
im.wkt <- geojson2wkt(toJSON(islamona$geojson$geometry))

#Save occurrence reports in the geometry, limited to preserved specimens
im.occ_results <- occ_search(geometry = im.wkt, basisOfRecord = 'PRESERVED_SPECIMEN')

#Save these results to a file
write.csv(im.occ_results$data, file = 'im.occ_results.csv')

#Plot results using sp and leaflet
library(sp)
library(leaflet)

s <- SpatialPoints(data.frame(cbind(im.occ_results$data$decimalLongitude, im.occ_results$data$decimalLatitude)))

map_leaf(s)
```
